Notes on $\etarho$

### 2009-02-05 docs/eta_notes/eta.tex
fixed a macro

The base state evolves in response to heating and mixing in the star.
The density evolution is governed by
\begin{eqnarray}
\frac{\partial \rho_0}{\partial t} &=& - 
 \nablab \cdotb \left( \rho_0 w_0 \er \right)
- \nablab \cdotb \left( \etarho \er \right) \enskip ,
\label{eq:rho0upd_new}
\end{eqnarray}
with
\begin{equation}
\etarho(r) = \overline{\left(\rhop \ut \cdot \er \right)} = \frac{1}{A(\Omega_H)} 
 \int_{\Omega_H}  (\rhop \ut \cdot \er ) \; dA \enskip , \label{eq:eta}
\end{equation}
designed to keep the average value of the full density, $\rho$, over a
layer of constant radius in the star equal to $\rho_0$.  To complete
the update of the base state, we need evolution equations for the
pressure, $p_0$, and velocity, $w_0$.  For spherical geometry, the
derivation of $w_0$ constraint equation is shown in
Appendix~\ref{app:gravity}, resulting in the following system
\begin{eqnarray}
w_0 &=& \ow + \delta w_0 \\
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar \\
\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=& - \frac{g}{r^2} \frac{\partial (r^2 \etarho)}{\partial r} - \frac{4 (\ow + \dw) \rho_0 g}{r} 
- 4 \pi G \rho_0 \etarho \label{eq:dw0constraint}
\end{eqnarray}

@@

\frac{\partial \rhop}{\partial t} = - \nablab \cdotb (\rho_0 \ut)
-\nablab \cdotb (\rhop w_0 \er) - \nablab \cdotb (\rhop \ut )

@@

\frac{\partial \rhop}{\partial t} + \nablab \cdotb (\rhop \ut) 
= - \nablab \cdotb (\rho_0 \ut)

@@

    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \ut ) \right) \; d\Omega 

@@

\int_{\Omega_H} \nabla \cdotb \left( \rho_0 \ut + \rhop w_0 \er \right) \; d\Omega \nonumber  \\

@@

\left [ \rho_0  (\ut \cdot \er)  + \rhop w_0 \right] r^2

@@

$\int_{\Omega_H} \rho_0 (\ut \cdot \er) d\Omega = 0$ follows from the definition of $\ut,$

@@

%\frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \int_{\Omega_H} \nabla \cdotb \left( \rho_0 \ut + \rhop w_0 \er \right)

@@

% \rhop (\ut \cdot \er)  \; r^2 d\Omega  \right  |_{r_0-h}^{r_0+h}

@@

    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \ut ) \right) \; d\Omega \nonumber \\

@@

  \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H}  \nabla \cdotb ( \rhop \ut ) \; d\Omega \right) \nonumber \\

@@

\left [ \int_{\Omega_H} \rhop (\ut \cdot \er) \; d\Omega  \right]  r^2\; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\

@@

\overline{[\rhop (\ut \cdot \er)]} \; r^2 \; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\

@@

\lim_{h\rightarrow 0} \frac{1}{r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \nabla \cdot \left[\overline{\left(\rhop\ut \cdot \er\right)} \er \right]\;  r^2 \; dr \nonumber \\
&=&  \frac{\partial}{\partial t} \overline{\rhop} + \nabla \cdot \left[ \overline{\left(\rhop \ut \cdot \er\right)} \er \right]

@@

In short, 
\begin{equation}
\frac{\partial}{\partial t} \overline{\rhop} = 
- \nabla \cdot \left [ \overline{\left(\rhop \ut \cdot \er\right)} \, \er \right ] = 
- \nabla \cdot ( \etarho \er )
 \enskip .\label{eq:rhopbar} 
\end{equation}

@@

Therefore, we take a different approach.  We compute $\etarho$ by
constructing the quantity $\rhop \ut \cdot \er$ in each cell, and then
use our average routine to construct a 1-d, cell-centered
$\eta_{\rho,r}$ (this is essentially numerically solving the integral
in Eq.~[\ref{eq:eta}]).  The edge-centered values of $\etarho$,
$\eta_{\rho,r+1/2}$ are then constructed by simple
averaging: \MarginPar{should we be doing a volume-weighted average?}
\begin{equation}
\eta_{\rho,r+1/2} = \frac{\eta_{\rho,r} + \eta_{\rho,r+1}}{2} \enskip .
\end{equation}

>>>>

The base state evolves in response to heating and mixing in the star.
The density evolution is governed by
\begin{eqnarray}
\frac{\partial \rho_0}{\partial t} &=& - 
 \nablab \cdotb \left( \rho_0 w_0 \er \right)
- \nablab \cdotb \left( \etarho \er \right) \enskip ,
\label{eq:rho0upd_new}
\end{eqnarray}
with
\begin{equation}
\etarho(r) = \overline{\left(\rhop \Ubt \cdot \er \right)} = \frac{1}{A(\Omega_H)} 
 \int_{\Omega_H}  (\rhop \Ubt \cdot \er ) \; dA \enskip , \label{eq:eta}
\end{equation}
designed to keep the average value of the full density, $\rho$, over a
layer of constant radius in the star equal to $\rho_0$.  To complete
the update of the base state, we need evolution equations for the
pressure, $p_0$, and velocity, $w_0$.  For spherical geometry, the
derivation of $w_0$ constraint equation is shown in
Appendix~\ref{app:gravity}, resulting in the following system
\begin{eqnarray}
w_0 &=& \ow + \delta w_0 \\
\frac{1}{r^2} \frac{\partial}{\partial r} \left (r^2 \ow \right ) &=& \Sbar \\
\frac{\partial}{\partial r} \left[ \frac{\gammabar p_0}{r^2} \frac{\partial}{\partial r} (r^2 \dw) \right] &=& - \frac{g}{r^2} \frac{\partial (r^2 \etarho)}{\partial r} - \frac{4 (\ow + \dw) \rho_0 g}{r} 
- 4 \pi G \rho_0 \etarho \label{eq:dw0constraint}
\end{eqnarray}

@@

\frac{\partial \rhop}{\partial t} = - \nablab \cdotb (\rho_0 \Ubt)
-\nablab \cdotb (\rhop w_0 \er) - \nablab \cdotb (\rhop \Ubt )

@@

\frac{\partial \rhop}{\partial t} + \nablab \cdotb (\rhop \Ubt) 
= - \nablab \cdotb (\rho_0 \Ubt)

@@

    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \Ubt ) \right) \; d\Omega 

@@

\int_{\Omega_H} \nabla \cdotb \left( \rho_0 \Ubt + \rhop w_0 \er \right) \; d\Omega \nonumber  \\

@@

\left [ \rho_0  (\Ubt \cdot \er)  + \rhop w_0 \right] r^2

@@

$\int_{\Omega_H} \rho_0 (\Ubt \cdot \er) d\Omega = 0$ follows from the definition of $\Ubt,$

@@

%\frac{1}{4\pi r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \int_{\Omega_H} \nabla \cdotb \left( \rho_0 \Ubt + \rhop w_0 \er \right)

@@

% \rhop (\Ubt \cdot \er)  \; r^2 d\Omega  \right  |_{r_0-h}^{r_0+h}

@@

    \left(\frac{\partial \rhop}{\partial t} + \nabla \cdotb ( \rhop \Ubt ) \right) \; d\Omega \nonumber \\

@@

  \int_{r_0-h}^{r_0+h} r^2 \; dr \int_{\Omega_H}  \nabla \cdotb ( \rhop \Ubt ) \; d\Omega \right) \nonumber \\

@@

\left [ \int_{\Omega_H} \rhop (\Ubt \cdot \er) \; d\Omega  \right]  r^2\; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\

@@

\overline{[\rhop (\Ubt \cdot \er)]} \; r^2 \; \right |_{r_0-h}^{r_0+h} \right) \nonumber  \\

@@

\lim_{h\rightarrow 0} \frac{1}{r_0^2 \, 2h} \int_{r_0-h}^{r_0+h} \nabla \cdot \left[\overline{\left(\rhop\Ubt \cdot \er\right)} \er \right]\;  r^2 \; dr \nonumber \\
&=&  \frac{\partial}{\partial t} \overline{\rhop} + \nabla \cdot \left[ \overline{\left(\rhop \Ubt \cdot \er\right)} \er \right]

@@

In short, 
\begin{equation}
\frac{\partial}{\partial t} \overline{\rhop} = 
- \nabla \cdot \left [ \overline{\left(\rhop \Ubt \cdot \er\right)} \, \er \right ] = 
- \nabla \cdot ( \etarho \er )
 \enskip .\label{eq:rhopbar} 
\end{equation}

@@

Therefore, we take a different approach.  We compute $\etarho$ by
constructing the quantity $\rhop \Ubt \cdot \er$ in each cell, and then
use our average routine to construct a 1-d, cell-centered
$\eta_{\rho,r}$ (this is essentially numerically solving the integral
in Eq.~[\ref{eq:eta}]).  The edge-centered values of $\etarho$,
$\eta_{\rho,r+1/2}$ are then constructed by simple
averaging: \MarginPar{should we be doing a volume-weighted average?}
\begin{equation}
\eta_{\rho,r+1/2} = \frac{\eta_{\rho,r} + \eta_{\rho,r+1}}{2} \enskip .
\end{equation}
